\Large \bf New Static Analysis Techniques to Detect Entropy Failure Vulnerabilities in Modern Software Projects

### 2018-12-14 paper/paper.tex
Added examples

We present inference rules for composing statements in the two programs $S_1 \cross S_1$ into one program that is semantically equivalent to $S_1\ ;\ S_2$. The most basic inference rule simply involves sequentially composing two atomic statements, which do not involve any control flow to be syncrhonized. Also, because the variables in the two versions of the program have been renamed to be disjoint, we can introduce an inference rule for the commutativity of two statements.

\begin{figure}
    \label{fig:infrules1}
    \caption{Basic Inference Rules}
    \[
		\begin{array}{c}
			\infer[]{A_1 \cross A_2 \leadsto A_1\ ;\ A_2}{} \\ \\ \\
			\infer[]{S_1 \cross S_2 \leadsto P}{
				S_2 \cross S_1 \leadsto P
			}
        \end{array}
    \]
\end{figure}

Next, we have the inference rule for synchronizing an if-statement with another statement. If $S_1$ is the taken branch, $S_2$ is the not-taken branch, and $S$ is the statement after the entire if-else-statement, then we would like to inject $S$ separately into each branch. This simple transformation is clearly equivalent to the sequential composition of the if-statement and $S$. The benefit of this transformation, however, is that the predicate abstraction tool has more information about the trace taken with which it can reason about product program.

Finally, we have the inference rule for synchronizing two while loops. The intuition here is that we would like the two loops to operate in lock-step as long as possible. To maintain the equivalence with the sequential composition, we also run the remaining iterations in the leftover loop after the other loop's condition no longer holds.

\begin{figure}
    \label{fig:infrules2}
    \caption{Conditional and Loop Inference Rules}
    \[
		\begin{array}{c}
			\infer[]{if(p)\ then\ S_1\ else\ S_2 \cross S \leadsto P}{
				\text{$S_1 \cross S \leadsto S_1'$} \\
				\text{$S_2 \cross S \leadsto S_2'$} \\
				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
			} \\ \\ \\
			\infer[]{while(p_1)\ S_1 \cross while(p_2)\ S_2 \leadsto P_0\ \cross\ P_1\ ;\ P_2}{
				\text{$P_0 = while(p_1 \land p_2)\ S_1\ ;\ S_2$} \\
				\text{$P_1 = while(p_1)\ S_1$} \\
				\text{$P_2 = while(p_2)\ S_2$}
			}
        \end{array}
    \]
\end{figure}

@@

@@

We define these ``unrelated'' statements to be the ones that are not tainted by any legitimate entropy source. This is reasonable because our ultimate assertions we would like to prove will only be over the tainted variables. Thus, we can use the $\Gamma$ environment we constructed in the instrumentation step to augment our inference rules. We say that $\Gamma \vdash S$ if some variable referenced in $S$ is tainted by at least one legitimate source of entropy. Similarly, $\Gamma \nvdash S$ if no variable in the statement is tainted by any legitimate source of entropy. We introduce an optimization inference rule that does not lead to any exponential blowup for statements that are not entailed by $\Gamma$.

\begin{figure}
    \label{fig:auginfrules1}
    \caption{Augmented Inference Rules}
    \[
		\begin{array}{c}
			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto S_1\ ;\ S_2}{
				\Gamma \nvdash S_1 \\
				\Gamma \nvdash S_2
			} \\
		\end{array}
    \]
\end{figure}

Then, we can make the corresponsding changes to our naive inference rules to recognize the $\Gamma$ environment we've introduced. This gives us the following set of augmented inference rules: % TODO: reference to figure

\begin{figure}
    \label{fig:auginfrules2}
    \caption{Augmented Inference Rules}
    \[
		\begin{array}{c}
			\infer[]{\Gamma \vdash A_1 \cross A_2 \leadsto A_1\ ;\ A_2}{} \\ \\ \\
			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto P}{
				\Gamma \vdash S_2 \cross S_1 \leadsto P
			} \\ \\ \\
			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto S_1\ ;\ S_2}{
				\Gamma \nvdash S_1 \\
				\Gamma \nvdash S_2
			} \\ \\ \\
			\infer[]{\Gamma \vdash if(p)\ then\ S_1\ else\ S_2 \cross S \leadsto P}{
				\text{$\Gamma \vdash S_1 \cross S \leadsto S_1'$} \\
				\text{$\Gamma \vdash S_2 \cross S \leadsto S_2'$} \\
				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
			} \\ \\ \\
			\infer[]{\Gamma \vdash while(p_1)\ S_1 \cross while(p_2)\ S_2 \leadsto P_0\ ;\ P_1\ ;\ P_2}{
				\text{$P_0 = while(p_1 \land p_2)\ S_1\ ;\ S_2$} \\
				\text{$P_1 = while(p_1)\ S_1$} \\
				\text{$P_2 = while(p_2)\ S_2$} \\
				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
			}
		\end{array}
    \]
\end{figure}

>>>>

We present inference rules for composing statements in the two programs $S_1 \cross S_1$ into one program that is semantically equivalent to $S_1\ ;\ S_2$. The most basic inference rule simply involves sequentially composing two atomic statements, which do not involve any control flow to be syncrhonized. Also, because the variables in the two versions of the program have been renamed to be disjoint, we can introduce an inference rule for the commutativity of two statements. See Figure \ref{fig:infrules1}.

\begin{figure}
    \caption{Basic Inference Rules}
    \label{fig:infrules1}
    \[
		\begin{array}{c}
			\infer[]{A_1 \cross A_2 \leadsto A_1\ ;\ A_2}{} \\ \\ \\
			\infer[]{S_1 \cross S_2 \leadsto P}{
				S_2 \cross S_1 \leadsto P
			}
        \end{array}
    \]
\end{figure}

Next, we have the inference rule for synchronizing an if-statement with another statement. If $S_1$ is the taken branch, $S_2$ is the not-taken branch, and $S$ is the statement after the entire if-else-statement, then we would like to inject $S$ separately into each branch. This simple transformation is clearly equivalent to the sequential composition of the if-statement and $S$. The benefit of this transformation, however, is that the predicate abstraction tool has more information about the trace taken with which it can reason about product program. See \ref{fig:infrules2}.

Finally, we have the inference rule for synchronizing two while loops. The intuition here is that we would like the two loops to operate in lock-step as long as possible. To maintain the equivalence with the sequential composition, we also run the remaining iterations in the leftover loop after the other loop's condition no longer holds. See \ref{fig:infrules2}.

\begin{figure}
    \caption{Conditional and Loop Inference Rules}
    \label{fig:infrules2}
    \[
		\begin{array}{c}
			\infer[]{if(p)\ then\ S_1\ else\ S_2 \cross S \leadsto P}{
				\text{$S_1 \cross S \leadsto S_1'$} \\
				\text{$S_2 \cross S \leadsto S_2'$} \\
				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
			} \\ \\ \\
			\infer[]{while(p_1)\ S_1 \cross while(p_2)\ S_2 \leadsto P_0\ \cross\ P_1\ ;\ P_2}{
				\text{$P_0 = while(p_1 \land p_2)\ S_1\ ;\ S_2$} \\
				\text{$P_1 = while(p_1)\ S_1$} \\
				\text{$P_2 = while(p_2)\ S_2$}
			}
        \end{array}
    \]
\end{figure}

@@

To demonstrate the application of these inference rules we will work an example. Consider the following two programs: Algorithm \ref{alg:p1} and \ref{alg:p2}.

\begin{algorithm}
  \caption{$P1$}
  \label{alg:p1}
  \begin{algorithmic}[1]
    \If {$p$}
        \State $x_1 \gets 1$
    \Else
        \State $x_1 \gets 2$
    \EndIf
  \end{algorithmic}
\end{algorithm}

and

\begin{algorithm}
  \caption{$P1$}
  \label{alg:p2}
  \begin{algorithmic}[1]
    \If {$p$}
        \State $x_1 \gets 2$
    \Else
        \State $x_1 \gets 1$
    \EndIf
  \end{algorithmic}
\end{algorithm}

These programs just switch the bodies of the if and else branches (so they are clearly not equivalent).

Sequentially composed, we could imagine passing the following program to CPAChecker: Algorithm \ref{alg:seq}.

\begin{algorithm}
  \caption{$P1; P2$}
  \label{alg:seq}
  \begin{algorithmic}[1]
    \If {$p$}
        \State $x_1 \gets 1$
    \Else
        \State $x_1 \gets 2$
    \EndIf
    \If {$p$}
        \State $x_1 \gets 2$
    \Else
        \State $x_1 \gets 1$
    \EndIf
    \State \textsf{assert}$(x_1 == x_2)$
  \end{algorithmic}
\end{algorithm}

But instead, we could synchronize the if statements with the if-statement inference rule to yield: Algorithm \ref{alg:cross}.

\begin{algorithm}
  \caption{$P1{} \cross P2$}
  \label{alg:cross}
  \begin{algorithmic}[1]
    \If {$p$}
        \State $x_1 \gets 1$
        \If {$p$}
            \State $x_1 \gets 2$
        \Else
            \State $x_1 \gets 1$
        \EndIf
    \Else
        \State $x_1 \gets 2$
        \If {$p$}
            \State $x_1 \gets 2$
        \Else
            \State $x_1 \gets 1$
        \EndIf
    \EndIf
    \State \textsf{assert}$(x_1 == x_2)$
  \end{algorithmic}
\end{algorithm}

Clearly Algorithm \ref{alg:cross} is twice as long because we replicated $P2$ into both branches of the if-statement. However this is a tradeoff we may be willing to make in order to ease the burden on our predicate abstraction tool as compared to Algorithm \ref{alg:seq}.

@@

We define these ``unrelated'' statements to be the ones that are not tainted by any legitimate entropy source. This is reasonable because our ultimate assertions we would like to prove will only be over the tainted variables. Thus, we can use the $\Gamma$ environment we constructed in the instrumentation step to augment our inference rules. We say that $\Gamma \vdash S$ if some variable referenced in $S$ is tainted by at least one legitimate source of entropy. Similarly, $\Gamma \nvdash S$ if no variable in the statement is tainted by any legitimate source of entropy. We introduce an optimization inference rule that does not lead to any exponential blowup for statements that are not entailed by $\Gamma$. See \ref{fig:auginfrules1}.

\begin{figure}
    \caption{Augmented Inference Rules}
    \label{fig:auginfrules1}
    \[
		\begin{array}{c}
			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto S_1\ ;\ S_2}{
				\Gamma \nvdash S_1 \\
				\Gamma \nvdash S_2
			} \\
		\end{array}
    \]
\end{figure}

Then, we can make the corresponding changes to our naive inference rules to recognize the $\Gamma$ environment we've introduced. This gives us the following set of augmented inference rules: See Figure \ref{fig:auginfrules2}.

\begin{figure}
    \caption{Augmented Inference Rules}
    \label{fig:auginfrules2}
    \[
		\begin{array}{c}
			\infer[]{\Gamma \vdash A_1 \cross A_2 \leadsto A_1\ ;\ A_2}{} \\ \\ \\
			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto P}{
				\Gamma \vdash S_2 \cross S_1 \leadsto P
			} \\ \\ \\
			\infer[]{\Gamma \vdash S_1 \cross S_2 \leadsto S_1\ ;\ S_2}{
				\Gamma \nvdash S_1 \\
				\Gamma \nvdash S_2
			} \\ \\ \\
			\infer[]{\Gamma \vdash if(p)\ then\ S_1\ else\ S_2 \cross S \leadsto P}{
				\text{$\Gamma \vdash S_1 \cross S \leadsto S_1'$} \\
				\text{$\Gamma \vdash S_2 \cross S \leadsto S_2'$} \\
				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
			} \\ \\ \\
			\infer[]{\Gamma \vdash while(p_1)\ S_1 \cross while(p_2)\ S_2 \leadsto P_0\ ;\ P_1\ ;\ P_2}{
				\text{$P_0 = while(p_1 \land p_2)\ S_1\ ;\ S_2$} \\
				\text{$P_1 = while(p_1)\ S_1$} \\
				\text{$P_2 = while(p_2)\ S_2$} \\
				\text{$P = if(p)\ then\ S_1'\ else\ S_2'$}
			}
		\end{array}
    \]
\end{figure}
